---
title: "Regression Models Course Project"
author: "Steven Zhang"
date: "21 Nov, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(GGally)
library(gridExtra)
```

# 1 - Synopsis

This report is about vehicles, aims to explore the relationship between miles per gallon(MPG) and some other variables. Particularly focused on the following questions:

- "Is an automatic or manual transmission better for MPG?"

- "Quantify the MPG difference between automatic and manual transmissions."

The original question description can be found [here](https://www.coursera.org/learn/regression-models/peer/nxntd/regression-models-course-project).

**From my analysis, the manual transmission car has a significant higher mpg than automatic's. In addition, from the generalized additive model, a manual transmission car has a fuel efficiency of 3.47 (MPG) higher than that of automatic transmission car.**

# 2 - Exploratory Analysis 

## 2.1 - Data Source Overview

The source dataset is embedded in R environment, named "mtcars", which was extracted from the 1974 *Motor Trend US* magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973¨C74 models). This dataset has 32 observations on 11 variables as followed. 

-	 mpg - Miles/(US) gallon
-	 cyl - Number of cylinders
-	 disp - Displacement (cu.in.)
-	 hp - Gross horsepower
-	 drat - Rear axle ratio
-	 wt - Weight (1000 lbs)
-	 qsec - 1/4 mile time
-	 vs - V/S
-	 am - Transmission (0 = automatic, 1 = manual)
-	 gear - Number of forward gears
-	 carb - Number of carburetors

## 2.2 - data loading and preprocessing

We load the data and convert the *am* variable to factor format.  

```{r loading}
df <- mtcars
summary(glm(mpg ~ ., data = df))$coefficients[,4]
```

From a simple analysis, I choose 5 variables for next steps.

```{r preprocessing}
df <- select(df, c(mpg, am, wt, qsec, hp, disp))
df$am <- as.factor(df$am)
```

Here is the correlation of the variables. For the convenience of further analysis, I hold the *Transmission* (auto/manual) as colors.

```{r exploratory1, warning=FALSE}
my.ggpairs <- function(data, mapping, ...){
  p <- ggplot(data = data, mapping = mapping, binwidth = 30) + 
    geom_point() + 
    geom_smooth(method=loess, fill="red", color="red", ...) +
    geom_smooth(method=lm, fill="blue", color="blue", ...)
  p
}
ggpairs(df, lower = list(continuous = my.ggpairs), aes(color = am))
```

From the diagram above, we can see some clear relationships between mpg and other variables. More, we can see a clear difference different transmissions (red & blue). 

# 3 - Analysis
## 3.1 - Question 1: Is an automatic or manual transmission better for MPG?

Now, we are going to explore the relationship between automatic and manual transmission. 

```{r exploratory2}
density.plot <- ggplot(df, aes(mpg, color = am)) + geom_density(legend = c("auto","manual"))
box.plot <- ggplot(df, aes(am, mpg)) + geom_boxplot(col = "skyblue4", legend = c("auto","manual"))
grid.arrange(density.plot, box.plot, ncol = 2, nrow = 1)
```

We can see from the plot above that there are obvious correlation between mpg and transmission. However, we should run a hypothesis test to confirm it. 

```{r t.test}
group.manual <- df[df$am == 1,]
group.auto <- df[df$am == 0,]
t.test(group.auto$mpg, group.manual$mpg, paired = FALSE, var.equal = FALSE)
```

From this T test, we can see the automatic transmission has a significant (p-value < 0.0015) less mpg value than manual's.

## 3.2 - Question 2: Quantify the MPG difference between automatic and manual transmissions

Based on the analysis above, we need to quantify the coefficients. I build three linear regression models with 1, 5 and 9 predictors and use the anova function to compare them.

```{r fit.models}
fit1 <- lm(mpg ~ am, data = df)
fit2 <- lm(mpg ~ ., data = df)
mtcars$am <- as.factor(mtcars$am)
fit3 <- lm(mpg ~ ., data = mtcars)
anova(fit1, fit2, fit3)
```

From this comparance, we choose model 2 to do further test. 

```{r summary.fit}
summary(fit2)
```

Based on this generalized additive model, we can say that a manual transmission car has a fuel efficiency of 3.47 (MPG) higher than that of a automatic transmission car. 

# 4 - Appendix

- [1] Henderson and Velleman (1981), Building multiple regression models interactively. *Biometrics*, 37, 391¨C411.

## 4.1 - Additional Plots

Here we list the plots of the regression.

```{r lm.plot}
par(mfrow = c(2, 2))
plot(fit2)
```

## 4.2 - Hardware & Software Environment

```{r env, echo=FALSE}
sessionInfo()
```